---
title: "Untitled"
author: "Venkata"
date: "June 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Data related to the execution of an exercise is provided.
In the data, there are 6 participants and 5 classe variables. The 5 classe variables are  
1.exactly according to the specification (Class A),   
2.throwing the elbows to the front (Class B),   
3.lifting the dumbbell only halfway (Class C),   
4.lowering the dumbbell only halfway (Class D)   
5.throwing the hips to the front (Class E).   

## GOAL: 
predict the manner in which the participants did the excercise.

### Step1:
Since participants have performed correctly and incorrectly, first step above all would be to look at
all the participants and classA-variable to see if there are any noticeable points to note.
### step2:
In the column new_window, there are only 2 values. 'yes' and 'no'. Considered the data with value = 'no' as there are more records with value 'no' (19216).
```{r include=FALSE}
library(caret);
library(ggplot2);
library(lattice);
setwd("C:/Users/Venkata/Desktop/Personal/Coursera/Data_Science_Specialization/Machine_Learning/assignment");
data <- read.csv("pml-training.csv")
#stats of the data.
colnames(data)
dim(data)
summary(data)
yes_window <- data[data$new_window=='yes',]
no_window <- data[data$new_window == 'no',]
data_subset <- no_window
dim(data_subset)
```

#Step3: 
Detemine what columns to use.
- Since the data could include both numeric and factor variables that could determine the performance of the model, for now, 
only the numeric data is considered.

```{r include=FALSE}
library(ggplot2)
colnames(data_subset)
listed_data <- as.list(data_subset)
column_class <- sapply(X= listed_data, FUN=class)
col_factor <- which(column_class=="factor")
numeric_columns <- which(column_class =="numeric")
temp_results <- sapply(X=data_subset[,numeric_columns], FUN=quantile, na.rm = TRUE)
dim(temp_results)
sub_data_subset <- numeric_columns[which(!is.na(temp_results[1,]))]
sub_data_subset
length(sub_data_subset)
useful_data_subset <- data_subset[,sub_data_subset]
colnames(useful_data_subset)
useful_data_subset$classe <- data_subset$classe
useful_data_subset$user_name <- data_subset$user_name
```
- I implemented cross-validation with cv-folds = 5 and cv-folds = 10. Keeping my computer's hardware in mind.
```{r include=FALSE}
library(randomForest)
library(caret)
inTrain <- createDataPartition(y=useful_data_subset$classe, p=0.75, list=FALSE)
inTrain
training <- useful_data_subset[inTrain,]
testing <- useful_data_subset[-inTrain,]
colnames(training)
res = cross_validation <- rfcv(trainx=training[,-26], trainy=training[,26], cv.fold=5, step=0.5, recursive=FALSE)
```